---
title: "R Notebook"
output: html_notebook
---



```{r}
library(ggplot2)
library(mgcv)
library(splines)
library(gam)
library(corrplot)
library("e1071")
library(GGally)
library(gbm)
library(caret)
library(factoextra)
library(MLmetrics)
library(xgboost)
library(caTools)
library(dplyr)
library(caret)

file_path <- "communities.data"

data <- read.table(file_path, sep = ",", header = TRUE, na.strings = "?")

head(data)
colnames(data)

```
```{r}
names_file <- "communities.names"
column_names <- readLines(names_file)

csv_file <- "communities.csv"

# Write the data to a .csv file
write.csv(data, file = csv_file, row.names = FALSE)

# Confirm the .csv file has been created and written correctly
print(paste("CSV file saved to", csv_file))
```

```{r}
# Provided string with column names
# Extract lines that start with '@attribute'
attribute_lines <- grep("^@attribute", column_names, value = TRUE)

# Extract column names from the attribute lines
column_names <- gsub("^@attribute\\s+([a-zA-Z0-9_]+)\\s+.*$", "\\1", attribute_lines)

# Print the extracted column names
print(column_names)

```
### Clean Data
```{r}
data_file <- "communities.data"

# Read the .data file into a data frame
data <- read.table(data_file, sep = ",", header = FALSE, na.strings = "?")

# Assign column names from the extracted list
colnames(data) <- column_names

# View the data to ensure the column names are correctly assigned
head(data)

# Read the .data file into a data frame
data <- read.table(data_file, sep = ",", header = FALSE, na.strings = "?")

# Assign column names from the extracted list
colnames(data) <- column_names

# View the data to ensure the column names are correctly assigned
head(data)
```
```{r}
null_prop <- colMeans(is.na(data))

# Set a threshold for proportion of NULL values
threshold <- 0.5  # For example, columns with over 50% NULL values will be removed

# Identify columns to remove
columns_to_remove <- names(null_prop[null_prop > threshold])

# Remove identified columns from the data frame
cleaned_data <- data[, !names(data) %in% columns_to_remove]

# View the cleaned data
print(cleaned_data)


cleaned_data<- na.omit(cleaned_data)

numeric_data <- select_if(cleaned_data, is.numeric)

# View the selected numeric variables
print(numeric_data)

pca_data <- numeric_data[, !names(numeric_data) %in% c("ViolentCrimesPerPop")]

# View the modified dataset
print(pca_data)

```
### PCA
```{r}
pca_result <- prcomp(pca_data, scale=TRUE)

plot(pca_result$x[,1:14], xlab="Principal Components", main="Principle Component Analysis" )
abline(h = 1, col = "blue", lty = 1) #eigenvalues should be > 1 

print(pca_result)
var <- get_pca_var(pca_result)

eigenvalues <- pca_result$sdev^2

# Identify PCs with eigenvalues greater than 1
pcs_gt_1 <- which(eigenvalues > 1)

# Filter PCA result and eigenvalues
pca_result_gt_1 <- pca_result
pca_result_gt_1$sdev <- pca_result$sdev[pcs_gt_1]
pca_result_gt_1$rotation <- pca_result$rotation[, pcs_gt_1]

# Plot PCA with PCs having eigenvalues greater than 1
plot(pca_result_gt_1, xlab = "Principal Components (Eigenvalue > 1)", main = "Principle Component Analysis (Eigenvalue > 1)")
abline(h = 1, col = "blue", lty = 1) # Eigenvalues should be > 1

# Extract variable contributions
var <- get_pca_var(pca_result_gt_1)
```


```{r}
eigenvalues <- pca_result$sdev^2
eigenvalues_subset <- eigenvalues[1:14]

pcs <- pca_result$x[, 1:14]

# Create a scatter plot matrix
pairs(pcs, main = "Scatter Plot Matrix of Principal Components")
```
### PCA plot

```{r}
variance_explained <- summary(pca_result)$importance[2, 1:20]  # Assuming you want to display the first 20 components

# Plot a grouped bar chart
barplot(variance_explained, 
        main = "Variance Explained by Principal Components",
        xlab = "Principal Component", ylab = "Variance Explained",
        col = "cornflowerblue", ylim = c(0, max(variance_explained) + 0.05),
        border = "black", names.arg = 1:20) + abline(h = 0.01, col = "red")
```

### ScreePlots
```{r}
prop_var_expl_sc = summary(pca_result)$importance[2,]
cum_var_expl_sc = summary(pca_result)$importance[3,]

plot(prop_var_expl_sc, type="b", xlab="Component number", ylab="Proportion of variance explained", ylim=c(0,1))
plot(cum_var_expl_sc, type="b", xlab="Number of components", ylab="Cumulative variance explained", ylim=c(0,1))

elbow_point <- which(diff(prop_var_expl_sc) < 0.1)[1]

# Fit a linear model (tangent) to the elbow point and the adjacent point
tangent_line <- lm(prop_var_expl_sc[elbow_point:(elbow_point + 1)] ~ c(elbow_point, elbow_point + 1))

# Plot the scree plot
plot(prop_var_expl_sc, type = "b", xlab = "Component number", ylab = "Proportion of variance explained", ylim = c(0, 1))

# Add the tangent line
abline(tangent_line, col = "red")
```


```{r}
library(pathviewr)

df <- data.frame(x = seq(1:102),
                 y = prop_var_expl_sc)
plot(df)
find_curve_elbow(df, plot_curve = TRUE)
```
```{r}
library(MAP)
# Now, you can call the MAP function
MAP::MAP(numeric_data)
```

### Horn's Parallel Analysis
```{r}
library(paran)
parallel_result <- paran(numeric_data)
```


```{r}
# Contributions of variables to PC1
fviz_contrib(pca_result, choice = "var", axes = 1, top = 10) + xlab("Variables (truncated)") + theme(axis.title.x = element_text(size = 14),axis.title.y = element_text(size = 14))
fviz_contrib(pca_result, choice = "var", axes = 2, top = 10) + xlab("Variables (truncated)") + theme(axis.title.x = element_text(size = 14), axis.title.y = element_text(size = 14))
fviz_contrib(pca_result, choice = "var", axes = 3, top = 10) + xlab("Variables (truncated)") + theme(axis.title.x = element_text(size = 14), axis.title.y = element_text(size = 14))
fviz_contrib(pca_result, choice = "var", axes = 4, top = 10) + xlab("Variables (truncated)") + theme(axis.title.x = element_text(size = 14),axis.title.y = element_text(size = 14))
fviz_contrib(pca_result, choice = "var", axes = 5, top = 10) + xlab("Variables (truncated)") + theme(axis.title.x = element_text(size = 14),axis.title.y = element_text(size = 14))
fviz_contrib(pca_result, choice = "var", axes = 6, top = 10) + xlab("Variables (truncated)") + theme(axis.title.x = element_text(size = 14),axis.title.y = element_text(size = 14))
fviz_contrib(pca_result, choice = "var", axes = 7, top = 10) + xlab("Variables (truncated)") + theme(axis.title.x = element_text(size = 14), axis.title.y = element_text(size = 14))
fviz_contrib(pca_result, choice = "var", axes = 8, top = 10) + xlab("Variables (truncated)") + theme(axis.title.x = element_text(size = 14), axis.title.y = element_text(size = 14))
fviz_contrib(pca_result, choice = "var", axes = 9, top = 10) + xlab("Variables (truncated)") + theme(axis.title.x = element_text(size = 14),axis.title.y = element_text(size = 14))
fviz_contrib(pca_result, choice = "var", axes = 10, top = 10) + xlab("Variables (truncated)") + theme(axis.title.x = element_text(size = 14),axis.title.y = element_text(size = 14))
fviz_contrib(pca_result, choice = "var", axes = 11, top = 10) + xlab("Variables (truncated)") + theme(axis.title.x = element_text(size = 14),axis.title.y = element_text(size = 14))
fviz_contrib(pca_result, choice = "var", axes = 12, top = 10) + xlab("Variables (truncated)") + theme(axis.title.x = element_text(size = 14),axis.title.y = element_text(size = 14))
fviz_contrib(pca_result, choice = "var", axes = 13, top = 10) + xlab("Variables (truncated)") + theme(axis.title.x = element_text(size = 14),axis.title.y = element_text(size = 14))
fviz_contrib(pca_result, choice = "var", axes = 14, top = 10) + xlab("Variables (truncated)") + theme(axis.title.x = element_text(size = 14),axis.title.y = element_text(size = 14))
```


```{r}

pca_scores <- pca_result$x[, 1:14]

new_data <- cbind(cleaned_data, pca_scores)

main_data_with_pca <- new_data[, -c(1:103)]
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.




### Gradient Boosting 
```{r}
n_rows <- nrow(main_data_with_pca)

# Generate indices for splitting
train <- sample(1:n_rows, 0.7 * n_rows)

# Split data into training and testing sets
train_data <- main_data_with_pca[train, ]
test_data <- main_data_with_pca[-train, ]
```


```{r}
mcr <- function(y, pred_prob){
  if (length(pred_prob) != length(y)){
    stop("Vectors should be the same length")
  } else {
    1 - sum(diag(table(y, pred_prob >= .5))) / length(pred_prob)
  }
}

#y_test <- 
#y_test <- as.numeric(MH_dat$dep_sev_fu)[test]-1
```

```{r}
grid <- expand.grid(shrinkage = c(.1, .01, .001),
                    n.trees = c(10, 100, 1000, 2000, 2500), 
                    interaction.depth = 1:4,
                    n.minobsinnode= 10)
```
```{r}
plot<-fviz_contrib(pca_result, choice = "ind", axes = 1:14)
plot + theme_minimal() + theme(axis.text.x = element_blank()) 
```

```{r}
set.seed(4168216)
gbmFit <- train(ViolentCrimesPerPop ~., data = main_data_with_pca[train, ], tuneGrid = grid, distribution="gaussian", method="gbm", trControl = trainControl(number=10L, verboseIter = TRUE)  )
```
```{r}
plot(gbmFit)
```

```{r}
gbmImp <- varImp(gbmFit, scale = FALSE)
gbmImp
```
```{r}
plot(gbmImp, top = 20)
```

```{r}
#importance <- summary(gbmFit, plotit = FALSE)
gbmFit$bestTune
gbmBest <- gbm(ViolentCrimesPerPop ~. , data = main_data_with_pca[train,], distribution="gaussian", n.trees= 1000, interaction.depth = 4, shrinkage = 0.01)
```

```{r}
pred_gbm = predict(gbmBest, newdata=main_data_with_pca[-train,]) 


mean_avg_error_gbm <- MAE(pred_gbm, main_data_with_pca[-train, ]$ViolentCrimesPerPop)
print(mean_avg_error_gbm)

```

```{r}
library(gbm)
library(pdp)

plot.gbm(gbmBest, i.var=1)
plot.gbm(gbmBest, i.var=2)
plot.gbm(gbmBest, i.var=3)
plot.gbm(gbmBest, i.var=4)
plot.gbm(gbmBest, i.var=5)
plot.gbm(gbmBest, i.var=6)
plot.gbm(gbmBest, i.var=7)
plot.gbm(gbmBest, i.var=8)
plot.gbm(gbmBest, i.var=9)
plot.gbm(gbmBest, i.var=10)
plot.gbm(gbmBest, i.var=11)
plot.gbm(gbmBest, i.var=12)
plot.gbm(gbmBest, i.var=13)
plot.gbm(gbmBest, i.var=14)
plot.gbm(gbmBest, i.var=15)
plot.gbm(gbmBest, i.var=16)
plot.gbm(gbmBest, i.var=17)
```

